### 2018-11-27 latex/discussion.tex
revised results again, changed naming into NamPT
revised supplementary table and implemented part of the comments from Alexander and Ines R.

We here comprehensively analysed the phylogenetic distribution of the three enzymes metabolising  Nam. These are the two NAD salvage pathway enzymes NADA and NamPRT as well as the Nam-degrading enzyme NNMT. We found that after the first appearance of NNMT in Protostomia, a diversification of NAD-consuming reactions  can be observed in Deuterostomia. We could explain these finding using mathematical modelling, as NNMT can remove excess Nam from cells and thereby reduce product inhibition of NAD signalling enzymes. This in turn enables higher fluxes through these reactions. Thus, the diversification of NAD-consuming enzymes in mammals might have been enabled by the presence of NNMT.

@@

We here developed a new approach that combines detailed phylogenetic analysis with dynamic metabolic modelling and have been able to explain observed evolutionary changes in the NAD biosynthesis and consumption pathway. Based on the simulated pathway dynamics, we have furthermore derived predictions for physiological interdependencies between several enzymes of the pathway that are potentially relevant for new disease treatments. Our results, including the experimental verification of parts of our predictions, demonstrate the potential of our approach for the analysis of dynamic networks and how the approach can be used to unravel functional interdependencies within pathways of interest.

>>>>

We here comprehensively analysed the phylogenetic distribution of the three enzymes metabolising  Nam. These are the two NAD salvage pathway enzymes NADA and NamPT as well as the Nam-degrading enzyme NNMT. We found that after the first appearance of NNMT in Protostomia, a diversification of NAD-consuming reactions  can be observed in Deuterostomia. We could explain these finding using mathematical modelling, as NNMT can remove excess Nam from cells and thereby reduce product inhibition of NAD signalling enzymes. This in turn enables higher fluxes through these reactions. Thus, the diversification of NAD-consuming enzymes in mammals might have been enabled by the presence of NNMT.

@@

We here developed an approach that combines detailed phylogenetic analysis with dynamic metabolic modelling and have been able to explain observed evolutionary changes in the NAD biosynthesis and consumption pathway. Based on the simulated pathway dynamics, we have furthermore derived predictions for physiological interdependencies between several enzymes of the pathway that are potentially relevant for new disease treatments. Our results, including the experimental verification of parts of our predictions, demonstrate the potential of our approach for the analysis of dynamic networks and how the approach can be used to unravel functional interdependencies within pathways of interest.

### 2018-11-27 latex/header.tex
revised supplementary table and implemented part of the comments from Alexander and Ines R.

\noindent
Mathias Bockwoldt\textsuperscript{1}, Dorothée Houry\textsuperscript{2}, Marc Niere\textsuperscript{2}, Toni I. Gossmann\textsuperscript{3}, Mathias Ziegler\textsuperscript{2}, and Ines Heiland\textsuperscript{1,§}

@@

>>>>

\noindent
Mathias Bockwoldt\textsuperscript{1}, Dorothée Houry\textsuperscript{2}, Marc Niere\textsuperscript{2}, Toni I. Gossmann\textsuperscript{3}, Ines Reinartz\textsuperscript{4,5}, Alexander Schug\textsuperscript{6}, Mathias Ziegler\textsuperscript{2}, and Ines Heiland\textsuperscript{1,§}

@@

\noindent
\textsuperscript{4}Department of Physics, Karlsruhe Institute of Technology, Wolfgang-Gaede-Str. 1, 76131 Karlsruhe, Germany
\textsuperscript{5}Steinbuch Centre for Computing, Karlsruhe Institute of Technology, Hermann-von-Helmholtz-Platz 1, 76344 Eggenstein-Leopoldshafen, Germany

\noindent
\textsuperscript{6}John von Neumann Institute for Computing, Jülich Supercomputing Centre, Forschungszentrum Jülich, 52425 Jülich, Germany

### 2018-11-27 latex/methods.tex
revised supplementary table and implemented part of the comments from Alexander and Ines R.

>>>>

\subsection{Molecular dynamics simulations}
We performed all-atom molecular dynamics simulations with explicit solvent for NamPT (PDB Code: 2H3D \cite{Wang2006}) with and without loop insertion. We used the AMBER99SB-ILDN force field \cite{Lindorff-Larsen2010} with the TIP3P water model \cite{Jorgensen1983} in GROMACS 5.1.2 \cite{Abraham2015}. The structures were simulated each in a box of water with distance between the solute and the box set to 0.2\,nm at a temperature of 300\,K for a total time of 1\,\textmu s. A time step of 2\,fs and the stochastic dynamics integrator were used. For the evaluation of the root mean square fluctuations (RMSF) the first 100\,ns of the simulations were omitted.

### 2018-11-27 latex/results.tex
revised supplementary table and implemented part of the comments from Alexander and Ines R.

The second hypothesis was  based on our model simulations that predict that NNMT might have exerted evolutionary pressure on  NamPT and that therefore the sequence insertion might have an effect on substrate binding of NamPT.  To analyse this possibility, we first compared different available protein structures of NamPT and tested whether  the loop insertion could potentially lead to dynamic structural rearrangements. This was done using homology modelling (see Figure ...) and molecular dynamics simulations for structures with and without the loop insertion. Taken together we were not able to show any obvious structural rearrangements. The molecular dynamics simulations showed little structural changes upon loop insertion and we observed a structurally stable catalytic core. This might be based on the fact that  all available protein structures of NamPT differ very little even at the catalytic side ( between 0.15Å and 0.4Å).

We nevertheless expressed and purified  both wildtype and mutant proteins, N-terminally fused to a 6xHis-tag in \textit{E.~coli}.. The size exclusion chromatography profile showed that both wildtype and mutant protein were expressed as dimers (see suppl. fig.~S3), indicating that the  mutant protein is likely to be folded correctly. The enzymatic activity was measured by NMR spectroscopy using the detection of NMN produced in the absence of ATP. Upon incubation with the NamPT inhibitor FK866 \cite{Hasmann2003} for 30 minutes, both wildtype and mutant NamPT did not produce any NMN, suggesting that binding of FK866 is not affected by the mutation (see suppl. fig.~S4). The mutant enzyme showed no detectable activity with 100 $\mu$M  Nam after 10 min of incubation at 30\textdegree{}C. With 1mM Nam ezymatic activity could be detected with the mutant enzyme. But the meassured activity is reduced to 30\% compared to the wildtype enzyme (fig.~\ref{fig:unresolved_loop}D). These observations suggest that the mutant enzyme is catalytically active, retains its dimeric state and sensitivity to FK866, but, has a  lower affinity to Nam, supporting the predictions derived from the metabolic modelling approach.

>>>>

The second hypothesis was  based on our model simulations that predict that NNMT might have exerted evolutionary pressure on  NamPT and that therefore the sequence insertion might have an effect on substrate binding of NamPT.  To analyse this possibility, we expressed and purified  both the wildtype and the mutant enzyme in \textit{E.~coli}., N-terminally fused to a 6xHis-tag. The size exclusion chromatography profile showed that both wildtype and mutant protein were expressed as dimers (see suppl. fig.~S3), indicating that the  mutant protein is likely to be folded correctly. The enzymatic activity was measured by NMR spectroscopy using the detection of NMN produced in the absence of ATP. Upon incubation with the NamPT inhibitor FK866 \cite{Hasmann2003} for 30 minutes, both wildtype and mutant NamPT did not produce any NMN, suggesting that binding of FK866 is not affected by the mutation (see suppl. fig.~S4). The mutant enzyme showed no detectable activity with 100 $\mu$M  Nam after 10 min of incubation at 30\textdegree{}C. With 1mM Nam enzymatic activity could be detected with the mutant enzyme. But the measured activity is reduced to 80\% of that observed for the wildtype enzyme under the same condition (fig.~\ref{fig:unresolved_loop}D). These observations suggest that the mutant enzyme is catalytically active, retains its dimeric state and sensitivity to FK866, but, has a  lower activity and affinity to Nam, supporting the predictions derived from the metabolic modelling approach.

To see whether we can find a mechanistic explanation for the reduced affinity of the mutant enzyme, we analysed different available protein structures of NamPT and tested whether  the loop insertion could potentially lead to dynamic structural rearrangements. To this end we applied homology modelling (fig.~\ref{fig:unresolved_loop}C) and molecular dynamics simulations for structures with and without the loop insertion (fig.~\ref{fig:unresolved_loop}F) Taken together we did not observe substantial structural rearrangements and the molecular dynamics simulations showed only limited structural changes upon loop insertion and we observed a mostly structurally stable catalytic core. This might be based on the fact that  all available protein structures of NamPT differ very little even at the catalytic side ( between 0.33Å and 0.95Å see supplementary table S3). Some residues close to the catalytic, nevertheless, showed slightly elevated mobilities in the wildtype structure, but, these elevated mobilities were dominated by rare events during the simulation time 1$~\mu s$. Thus, they do not appear as a robust change of structural dynamics upon loop insertion. They could nevertheless provide a hind towards reduced mobility of the mutant enzyme that explain the observed changes in affinity and activity upon loop deletion.
